Using reference models in feature selection problems \vspace{.1in}

### 2019-11-08 paper/ref_approach.tex
fixed TODO and plots

\documentclass[american,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The reference model approach in feature selection problems},
            pdfauthor={Federico Pavone, Juho Piironen, Paul-Christian B\"{u}rkner, Aki Vehtari},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=RoyalBlue,
            filecolor=black,
            urlcolor=RoyalBlue,
          }
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=american]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\fi
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{placeins}
\usepackage{graphicx,grffile,subcaption}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{3}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

Figure \ref{fig:entropy} reports the entropy scores for the different
methods compared. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy suffers of such a problem, thus
such a measure should be seen as one of different possible
measurements of the actual stability. 
[TODO: please clarify the above sentences about entropy.] 
The use of a reference model
improves slightly stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

We can use the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,..,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \}
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule.
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.
\hl{NOTE: if necessary I can add a pseudo-code illustration of the iterative projection algorithm, as 
I did in the thesis}
TODO: add the pseudo code here

@@

Figure \ref{fig:sensitivity_vs_fdr_iterated} reports the average sensitivity on
the vertical axis versus the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. For each method (different colours)
the squares and the circles correspond to using the reference and
approach respectively. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger number of
observations and the higher the true correlations, the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant covariates. The iterative
projection performs fairly well if the number of observations and the correlation level
are not too small. In those cases it mainly results to be too conservative, resulting in a low
sensitivity, but good false discovery rate. Tuning $\alpha$ might lead to improved selection, yet
the main source of poor performance could well be the miscalibration of the uncertainty of the predictive
utility score for misspecified models with cross-validation \hl{(maybe add some references?)}.

>>>>

\documentclass[american,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\usepackage{algorithm2e}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The reference model approach in feature selection problems},
            pdfauthor={Federico Pavone, Juho Piironen, Paul-Christian B\"{u}rkner, Aki Vehtari},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=RoyalBlue,
            filecolor=black,
            urlcolor=RoyalBlue,
          }
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=american]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\fi
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{placeins}
\usepackage{graphicx,grffile,subcaption}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{3}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

Figure \ref{fig:entropy} reports the entropy scores for the different
methods compared. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy can not distinguish the interchangeability 
due to correlation from instability, thus such a measure should be 
consider as relative indicator and not as an absolute measure of stability.
\hl{TODO: The above sentence should be clearer now}
The use of a reference model
improves slightly stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

We can use the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,..,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \}
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.

\begin{center}
\begin{algorithm}[H]
\SetAlgoLined
\KwResult{$R:=\{\text{selected features}\}$}
 $F:=\{\text{set of features}\}$ \;
 $R=\{\emptyset\}$ \;
 Fit reference model\;
 \While{$F\neq\{\emptyset\}$}{
   projection.HeuristicSearch \;
   projection.elpdEstimate() \;
%:
   $S=\text{min}\{\text{sub}: P(\text{elpd}_{\text{sub}}-\text{elpd}_{\text{base}}>0)\geq\alpha)$\} \;

  \eIf{$S=\{\emptyset\}$}{
   break \;
   }{
   $R=R\cup S$\;
   $F=F\setminus S$\;
  }
 }
 \caption{Automated iterative projections}
 \label{alg:iterated_proj}
\end{algorithm}
\end{center}

@@

Figure \ref{fig:sensitivity_vs_fdr_iterated} reports the average sensitivity on
the vertical axis versus the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. For each method (different colours)
the squares and the circles correspond to using the reference and
approach respectively. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger number of
observations and the higher the true correlations, the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant covariates. The iterative
projection performs fairly well if the number of observations and the correlation level
are not too small. In those cases it mainly results to be too conservative, resulting in a low
sensitivity, but good false discovery rate. Tuning $\alpha$ might lead to improved selection, yet
the main source of poor performance could well be the miscalibration of the uncertainty of the predictive
utility score for misspecified models with cross-validation \hl{(maybe add some references? The last sentence 
refers to the problem when comparing misspecified models using CV. Tuomas did a talk about this problem.
Maybe Aki can check the content and add some references - not sure Tuomas' work is already available)}.
